---
title: "King County Traverse: 2023"
output: 
  html_document:
    includes:
      after_body: footer.html
editor_options: 
  chunk_output_type: console
---

<link rel="stylesheet" href="kraustralia/kraustralia.css" type="text/css">

<br/>
<br/>
<br/>

I like maps.  I like doing fun things with spatial data in R.  Recently, I've taken up running and, when given the chance, trail running.  Running through the woods is my Zen.  

I prefer new trails to ones I've done before.  To push me to explore more trails, I took on the challenge of running across King Count (East-West direction) in 2023. From the Lighthouse at Discovery Park to the warming shed at Snoqualmie Pass/Hyak. And to do so by taking trails whenever possible, with a priority for non-paved.  Roads were used only when no reasonable trail option existed (unpaved Forest Service Road count as trails). This was made a lot easier by the [Palouse to Cascades State Park Trail](https://parks.wa.gov/find-parks/state-parks/palouse-cascades-state-park-trail).  

It wasn't always fun or easy.  There are some odd, ugly and difficult trails needed to tie all this together.  I got lost a few times despite my GPS.  I ran through some snow, some rain and a the occasional overgrown trail full of devil's club.  I saw two Elk, many vestiges of last century's industrial prowess and -- other than a few places on Cougar, Tiger, Rattlesnake and my city runs -- relatively few other runners.  

99 Miles, 85 of them on trails.  Many hours of fun in they woods and some in the city. 

Next up? North-South? Kittitas County?  Washington?  

```{r, echo = FALSE, message = FALSE, warning=FALSE, comment = FALSE}
 ## Load libraries
  library(sf)
  library(geosphere)
  library(tidyverse)
  library(basemaps)
  library(maps)
  library(tmap)

 ## Source Custom Functions
  source("~/projects/andykrause.github.io/functions/mapping_functions.R")

 ## Set Paths

  file_path <- '~/documents/_travel/america'
```

```{r, echo = FALSE, message = FALSE, warning=FALSE, comment = FALSE}

  ## Load Running Shapefile
  running_sf <- sf::st_read(file.path(file_path, 'running.shp'),
                            quiet=TRUE) %>%
    sf::st_transform(4269) 

  traverse_sf <- sf::st_read(file.path(file_path, '_events', 'king_county_traverse.shp'),
                            quiet=TRUE) %>%
    sf::st_transform(4269) 
  
   boundary_sf <- sf::st_read(file.path(file_path, '_events', 'king_county_boundary.shp'),
                            quiet=TRUE) %>%
    sf::st_transform(4269) 
  
  ## Split into trail/road 

  trailruns_sf <- running_sf %>%
    dplyr::filter(!is.na(type)) %>%
    dplyr::filter(type == 'trail') %>%
    dplyr::select(name, type) %>%
    group_by(name, type) %>% 
    summarise(segments = dplyr::n()) %>% 
    st_cast() %>%
    dplyr::select(-segments)

  roadruns_sf <- running_sf %>%
    dplyr::filter(!is.na(type)) %>%
    dplyr::filter(type != 'trail') %>%
    dplyr::select(name, type)
  
  trailruns_tf <- traverse_sf %>%
    dplyr::filter(!is.na(type)) %>%
    dplyr::filter(type == 'trail') %>%
    dplyr::select(name, type) %>%
    group_by(name, type) %>% 
    summarise(segments = dplyr::n()) %>% 
    st_cast() %>%
    dplyr::select(-segments)
  
  roadruns_tf <- traverse_sf %>%
    dplyr::filter(!is.na(type)) %>%
    dplyr::filter(type != 'trail') %>%
    dplyr::select(name, type) %>%
    group_by(name, type) %>% 
    summarise(segments = dplyr::n()) %>% 
    st_cast() %>%
    dplyr::select(-segments)
  

  `Trail Runs` <- trailruns_sf
  `Road Runs` <- roadruns_sf
  `King County Traverse: Trail` <- trailruns_tf
  `King County Traverse: Road` <- roadruns_tf
  `King County` <- boundary_sf

  trail_col <- 'yellow'
  road_col <- 'burlywood4'
  basemap <- leaflet::providers$CartoDB.DarkMatter
  
  ttable <- traverse_sf %>%
    dplyr::group_by(type) %>%
    dplyr::summarize(dist = sum(Length/1600))
  
```

&nbsp;
&nbsp;

```{r, echo = FALSE, cache = FALSE, comment = FALSE, warning = FALSE, message = FALSE, fig.width = 8, fig.height = 5.5}

## King County Traverse

bbox_sea <- st_bbox(trailruns_sf) # current bounding box
bbox_sea[1] <- -122.45
bbox_sea[2] <- 47.26
bbox_sea[3] <- -121.67
bbox_sea[4] <- 47.77

tmap_mode("view")

tm_shape(`Trail Runs`, bbox = bbox_sea) + 
   tm_basemap(basemap) +
   tm_lines(col = trail_col, scale = 1, legend.lwd.show = TRUE, alpha=.2) +
   tm_shape(`Road Runs`, bbox = bbox_sea) + 
     tm_lines(col = road_col, scale = 1, legend.lwd.show = FALSE, alpha=.2)+
   tm_shape(`King County Traverse: Trail`, bbox = bbox_sea) + 
     tm_lines(col = 'red', scale = 2, legend.lwd.show = FALSE)+
   tm_shape(`King County Traverse: Road`, bbox = bbox_sea) + 
     tm_lines(col = 'indianred4', scale = 2, legend.lwd.show = FALSE)+
   tm_shape(`King County`, bbox = bbox_sea) + 
     tm_borders(col = 'gray50')

```
 