---
title: "King County Traverse: 2023"
output: 
  html_document:
    includes:
      after_body: footer.html
editor_options: 
  chunk_output_type: console
---

<link rel="stylesheet" href="kraustralia/kraustralia.css" type="text/css">

<br/>
<br/>
<br/>

I like maps.  I also like doing fun things with spatial data in R.  Recently, I've taken up running and, when given the chance, trail running.  Running through the woods is my zen.  

I prefer new trails to ones I've done before. To push me to explore more trails I took on the challenge of running across King County (East-West direction) in 2023. From the Lighthouse at Discovery Park to the warming shed at Snoqualmie Pass/Hyak. And to do so by taking trails whenever possible, with a priority for non-paved.  Roads were used only when no reasonable trail option existed (I counted unpaved Forest Service Road as trails). This was made a lot easier by the [Palouse to Cascades State Park Trail](https://parks.wa.gov/find-parks/state-parks/palouse-cascades-state-park-trail).  

It wasn't always fun or easy.  There are some odd, ugly and difficult trails needed to tie all this together.  I got lost a few times despite my GPS.  I ran through some snow, some rain and the occasional overgrown thicket of devil's club.  Out on the trails I saw two elk, a coyote, many vestiges of last century's industrial prowess and -- other than a few places on Cougar and Rattlesnake -- relatively few other runners.  

105 Miles, 94 of them on trails. Intentionally not the most direct path, but many hours of fun in the woods and some in the city. 

Next up? North-South? Kittitas County?  Washington?  

```{r, echo = FALSE, message = FALSE, warning=FALSE, comment = FALSE}
 ## Load libraries
  library(sf)
  library(geosphere)
  library(tidyverse)
  library(basemaps)
  library(maps)
  library(tmap)

 ## Source Custom Functions
  source(file.path(getwd(), "/functions/mapping_functions.R"))

 ## Set Paths

 data_path <- '~/dropbox/andy/data'
```

```{r, echo = FALSE, message = FALSE, warning=FALSE, comment = FALSE}

  ## Load Running Shapefile
  running_sf <- sf::st_read(file.path(data_path, 'where', 'america', 'running.shp'),
                            quiet=TRUE) %>%
    sf::st_transform(4269) 

  traverse_sf <- sf::st_read(file.path(data_path, 'where', 'events', 'king_county_traverse.shp'),
                            quiet=TRUE) %>%
    sf::st_transform(4269) 
  
   boundary_sf <- sf::st_read(file.path(data_path, 'base', 'administrative', 'king_county_boundary.shp'),
                            quiet=TRUE) %>%
    sf::st_transform(4269) 
  
  ## Split into trail/road 

  trailruns_sf <- running_sf %>%
    dplyr::filter(!is.na(type)) %>%
    dplyr::filter(type == 'trail') %>%
    dplyr::select(name, type) %>%
    group_by(name, type) %>% 
    summarise(segments = dplyr::n()) %>% 
    st_cast() %>%
    dplyr::select(-segments)

  roadruns_sf <- running_sf %>%
    dplyr::filter(!is.na(type)) %>%
    dplyr::filter(type != 'trail') %>%
    dplyr::select(name, type)
  
  trailruns_tf <- traverse_sf %>%
    dplyr::filter(!is.na(type)) %>%
    dplyr::filter(type == 'trail') %>%
    dplyr::select(name, type) %>%
    group_by(name, type) %>% 
    summarise(segments = dplyr::n()) %>% 
    st_cast() %>%
    dplyr::select(-segments)
  
  roadruns_tf <- traverse_sf %>%
    dplyr::filter(!is.na(type)) %>%
    dplyr::filter(type != 'trail') %>%
    dplyr::select(name, type) %>%
    group_by(name, type) %>% 
    summarise(segments = dplyr::n()) %>% 
    st_cast() %>%
    dplyr::select(-segments)
  

  `Trail Runs` <- trailruns_sf
  `Road Runs` <- roadruns_sf
  `King County Traverse: Trail` <- trailruns_tf
  `King County Traverse: Road` <- roadruns_tf
  `King County` <- boundary_sf

  trail_col <- 'yellow'
  road_col <- 'burlywood4'
  basemap <- leaflet::providers$CartoDB.DarkMatter
  
  ttable <- traverse_sf %>%
    dplyr::group_by(type) %>%
    dplyr::summarize(dist = sum(Length/1600))
  
```

&nbsp;
&nbsp;

```{r, echo = FALSE, cache = FALSE, comment = FALSE, warning = FALSE, message = FALSE, fig.width = 8, fig.height = 5.5}

## King County Traverse

bbox_sea <- st_bbox(trailruns_sf) # current bounding box
bbox_sea[1] <- -122.45
bbox_sea[2] <- 47.26
bbox_sea[3] <- -121.67
bbox_sea[4] <- 47.77

tmap_mode("view")

tm_shape(`Trail Runs`, bbox = bbox_sea) + 
   tm_basemap(basemap) +
   tm_lines(col = trail_col, lwd = 1, col_alpha=.2) +
   tm_shape(`Road Runs`, bbox = bbox_sea) + 
     tm_lines(col = road_col, lwd = 1, col_alpha=.2)+
   tm_shape(`King County Traverse: Trail`, bbox = bbox_sea) + 
     tm_lines(col = 'red', lwd = 2)+
   tm_shape(`King County Traverse: Road`, bbox = bbox_sea) + 
     tm_lines(col = 'indianred4', lwd = 2)+
   tm_shape(`King County`, bbox = bbox_sea) + 
     tm_borders(col = 'gray50')

```
 